Learning From Massive Noisy Labeled Data for Image Classification

### 2015-04-10 paper.tex
Polish the paper again

Deep learning with a massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2014panda}, and scene classification~\cite{farabet2013learning}. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

@@

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels~\cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}. Mnih and Hinton~\cite{mnih2012learning} built a simple noise model for aerial images but only considered binary classification. Larsen \etal~\cite{larsen1998design} assumed label noises are independent from true class labels which is a simple and special case. Sukhbaatar \etal~\cite{sukhbaatar2014learning} generalized from them by considering multi-class classification and modeling class dependent noise, but they assumed the noise was conditionally independent with the image content, ignoring the hardness of labeling images of different confusing levels. Our work can be viewed as a generalization of~\cite{sukhbaatar2014learning,natarajan2013learning} and our model is flexible enough to not only class dependent noise but also input dependent noise.
% section related_work (end)

@@

\subsection{Learning the Parameters} % (fold)
\label{sub:learning_the_parameters}
We exploit two CNNs to model $p(\vec{y} | \vec{x})$ and $p(\vec{z} | \vec{x})$ separately. Denote the parameter set of each CNN by $\theta_1$ and $\theta_2$. Our goal is to find the optimal $\theta = \theta_1 \cup \theta_2$ that maximize the incomplete log-likelihood $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$. The EM algorithm is used to iteratively solve this problem.

@@

After both CNN components are properly pretrained, we can start to train the whole network with massive noisy labeled data. However, some practical issues need to be further discussed. First, if we merely use noisy labels, we will lose precious knowledge that we have gained before and the model could be drifted. Therefore, we need to mix the data with clean labels into our training set, which is depicted in Figure~\ref{fig:framework} as the extra supervisions for the two CNNs. Then each CNN receives two kinds of gradients, one is from the clean labels and the other is from the noisy labels. We denote them by $\Delta_c$ and $\Delta_n$, respectively. A potential problem is that $|\Delta_c| \ll |\Delta_n|$, because clean data is much less than noisy data. To deal with this problem, we bootstrap the clean data $\mathcal{D}_c$ to half amount of the noisy data $\mathcal{D}_{\eta}$. This upsampling process brings another advantage --- the gradients we calculated in each mini-batch are much more stable.

Our proposed method has the ability to figure out the true label given the image and its noisy label. From the perspective of information, our model predicts from two kinds of clues: what are the true labels for other similar images; and how easily for the image to be mislabeled. Label Propagation method~\cite{zhu2002learning} explicitly uses the first kind of information, while we implicitly capture it with a discriminative deep model. On the other hand, the second kind of information correlates the image content with the label noise, which can help distinguish between hard samples and mislabeled ones.
% section deep_learning_from_noisy_labels (end)

@@

\begin{table*}[t]
\begin{center}
\begin{tabular}{c|c|c|c|c}
\hline
\# & Method & Training Data & Initialization & Test Accuracy\\
\hline\hline
1 & AlexNet & $\mathcal{D}_c$ & random & 64.54\% \\
2 & AlexNet & $\mathcal{D}_c$ & ImageNet pretrained & 72.63\% \\
3 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & random & 74.03\% \\
4 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & ImaegNet pretrained & 75.13\% \\
5 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & model \#2 & 75.30\% \\
6 & Pseudo-Label\cite{lee2013pseudo} & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as unlabeled & model \#2 & 73.04\% \\
7 & Bottom-Up~\cite{sukhbaatar2014learning} & $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ & model \#2 & 76.22\% \\
8 & Ours & $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ & model \#2 & \textbf{78.24\%} \\
\hline
\end{tabular}
\end{center}
\caption{Experiment results on the collected dataset}
\label{tab:exp}
\end{table*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/model_output.pdf}
\caption{Examples of handling noisy labels. The information layout in each block is illustrated in the top-left one.}
\label{fig:model_output}
\end{center}
\end{figure*}

@@

Next, from row \#6 we can see that semi-supervised learning methods may not be a good choice when massive noisy labeled data are available. Although model \#6 achieves marginally better result than its base model, it is significantly inferior to model \#5, which indicates that simply discarding all the noisy labels cannot make the full use of these information.

@@

Next we demonstrate the effect of learning to predict the label noise type. We estimate $p(\vec{z}_2=1|\vec{x})+p(\vec{z}_3=1|\vec{x})$ on the validation set and sort the images accordingly in descending order. Then for each image, we check whether its noisy label mismatches its clean label. The rank-precision curve is plotted in Figure~\ref{fig:noise_prediction_rp}. It shows that the intuitive observation --- mislabeled images often share similar patterns --- is reasonable, and our model is trained to recognize such patterns to help handle noisy labels.

@@

\section{Conclusion} % (fold)
\label{sec:conclusion}
In this paper, we raised the problem of training a classifier with limited clean annotations and massive noisy labeled data. We proposed a novel probabilistic model to describe how a noisy label is generated. Two latent variables --- true label and noise type --- were introduced to bridge the semantic gap between the observed image and its noisy label. We solved the problem by the EM algorithm and integrated it into a deep learning framework. Experiments on a collected large-scale clothing dataset showed that massive noisy label data could benefit the training of deep models, and utilizing our method could further improve the performance.
% section conclusion (end)

>>>>

Deep learning from large-scale supervised training dataset has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2014panda}, and scene classification~\cite{farabet2013learning}. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

@@

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels~\cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}. Mnih and Hinton~\cite{mnih2012learning} built a simple noise model for aerial images but only considered binary classification. Larsen \etal~\cite{larsen1998design} assumed label noises are independent from true class labels which is a simple and special case. Sukhbaatar \etal~\cite{sukhbaatar2014learning} generalized from them by considering multi-class classification and modeling class dependent noise, but they assumed the noise was conditionally independent with the image content, ignoring the hardness of labeling images of different confusing levels. Our work can be viewed as a generalization of~\cite{sukhbaatar2014learning,natarajan2013learning} and our model is flexible enough to not only class dependent but also image dependent noise.
% section related_work (end)

@@

\subsection{Learning the Parameters} % (fold)
\label{sub:learning_the_parameters}
We exploit two CNNs to model $p(\vec{y} | \vec{x})$ and $p(\vec{z} | \vec{x})$ separately. Denote the parameter set of each CNN by $\theta_1$ and $\theta_2$. Our goal is to find the optimal $\theta = \theta_1 \cup \theta_2$ that maximizes the incomplete log-likelihood $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$. The EM algorithm is used to iteratively solve this problem.

@@

After both CNN components are properly pretrained, we can start to train the whole network with massive noisy labeled data. However, some practical issues need further discussion. First, if we merely use noisy labels, we will lose precious knowledge that we have gained before and the model could be drifted. Therefore, we need to mix the data with clean labels into our training set, which is depicted in Figure~\ref{fig:framework} as the extra supervisions for the two CNNs. Then each CNN receives two kinds of gradients, one is from the clean labels and the other is from the noisy labels. We denote them by $\Delta_c$ and $\Delta_n$, respectively. A potential problem is that $|\Delta_c| \ll |\Delta_n|$, because clean data is much less than noisy data. To deal with this problem, we bootstrap the clean data $\mathcal{D}_c$ to half amount of the noisy data $\mathcal{D}_{\eta}$. In our experiments, we find that the performance of the classifier drops significantly without upsampling, but it is not sensitive with the upsampling ratio as long as the number of clean and noisy samples are in the same order.

Our proposed method has the ability to figure out the true label given the image and its noisy label. From the perspective of information, our model predicts from two kinds of clues: what are the true labels for other similar images; and how likely the image is mislabeled. The Label Propagation method~\cite{zhu2002learning} explicitly uses the first kind of information, while we implicitly capture it with a discriminative deep model. On the other hand, the second kind of information correlates the image content with the label noise, which can help distinguish between hard samples and mislabeled ones.
% section deep_learning_from_noisy_labels (end)

@@

\begin{table*}[t]
\begin{center}
\begin{tabular}{c|c|c|c|c}
\hline
\# & Method & Training Data & Initialization & Test Accuracy\\
\hline\hline
1 & AlexNet & $\mathcal{D}_c$ & random & 64.54\% \\
2 & AlexNet & $\mathcal{D}_c$ & ImageNet pretrained & 72.63\% \\
3 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & random & 74.03\% \\
4 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & ImaegNet pretrained & 75.13\% \\
5 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & model \#2 & 75.30\% \\
6 & Pseudo-Label\cite{lee2013pseudo} & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as unlabeled & model \#2 & 73.04\% \\
7 & Bottom-Up~\cite{sukhbaatar2014learning} & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ & model \#2 & 76.22\% \\
8 & Ours & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ & model \#2 & \textbf{78.24\%} \\
\hline
\end{tabular}
\end{center}
\caption{Experiment results on the collected dataset}
\label{tab:exp}
\end{table*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\linewidth]{figure/model_output.pdf}
\caption{Examples of handling noisy labels. The information layout in each block is illustrated in the top-left one.}
\label{fig:model_output}
\end{center}
\end{figure*}

@@

Next, from row \#6 we can see that semi-supervised learning methods may not be a good choice when massive noisy labeled data are available. Although model \#6 achieves marginally better result than its base model, it is significantly inferior to model \#5, which indicates that simply discarding all the noisy labels cannot fully utilize these information.

@@

Next we demonstrate the effect of learning to predict the label noise type. We first estimate $p(\vec{z}_2=1|\vec{x})+p(\vec{z}_3=1|\vec{x})$ on the validation set and sort the images accordingly in descending order. Then for each image, we check whether its noisy label mismatches its clean label. The rank-precision curve is plotted in Figure~\ref{fig:noise_prediction_rp}. It shows that the intuitive observation --- mislabeled images often share similar patterns --- is reasonable, and our model is trained to recognize such patterns to help handle noisy labels.

@@

\section{Conclusion} % (fold)
\label{sec:conclusion}
In this paper, we raised the problem of training a classifier with limited clean annotations and massive noisy labeled data. We proposed a novel probabilistic model to describe how a noisy label is generated. Two latent variables --- true label and noise type --- were introduced to bridge the semantic gap between an observed image and its noisy label. We solved the problem by the EM algorithm and integrated it into a deep learning framework. Experiments on a collected large-scale clothing dataset showed that massive noisy label data could benefit the training of deep models, and utilizing our method could further improve the performance.
% section conclusion (end)
